---
title: "Code HW4"
output: pdf_document
---
Loading the required libraries
```{r} 
require(gdata)
library(readxl)
library('reshape2')
library(gplots)
library(ggplot2)
library(moments)
library(dplyr)
```
Loading in the File
```{r}
gauge <- read_excel("~/Desktop/Quarter12/MATH189/HW4/gauge.xlsx")
```

PART I:

Simple linear regression
```{r}
model <- lm(gauge$density ~ gauge$gain)
yfit<-model$fitted.values
summary(lm(gauge$density ~ gauge$gain))
plot(gauge$gain,gauge$density,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')
lines(gauge$gain, yfit)
```

Creating the weights, with variance
```{r}
d <- gauge$gain
x=0
store=vector()
for (i in 1:9){
  variance=vector()
  for (j in 1:10){
    variance[j] <- d[j+x]
  }
  x=x+10
  store[i] <- var(variance)
}
store <- rep(store,each=10)
```

Creating the means for later
```{r}
gauge <- read_excel("~/Desktop/Quarter12/MATH189/HW4/gauge.xlsx")
g <- gauge$gain
x=0
newgain=vector()
for (i in 1:9){
  avg=vector()
  for (j in 1:10){
    avg[j] <- g[j+x]
  }
  x=x+10
  newgain[i] <- mean(avg)
}

d <- gauge$density
y=0
newdens=vector()
for (k in 1:9){
  newdens[k] <- d[k+y]
  y=y+10
}
```

Simple linear regression
```{r}
model <- lm(newdens ~ newgain)
yfit<-model$fitted.values
summary(lm(newdens ~ newgain))
plot(newgain,newdens,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')
lines(newgain, yfit)
```

QQPlot Linear Regression

```{r}
d <- gauge
fit <- lm(density ~ gain, data=d)
head(fit, 20)
summary(fit)
plot(fit)
sprintf("Kurtosis of Residuals: %f", kurtosis(fit))
sprintf("Skewness of Residuals: %f", skewness(fit))
d$predicted <- predict(fit)   # Save the predicted values
d$residuals <- residuals(fit) # Save the residual values

```

```{r}
d %>% select(density, predicted, residuals) %>% head()
rp <- ggplot(d, aes(x = gain, y = density)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightblue") +  # Plot regression slope
  geom_segment(aes(xend = gain, yend = predicted), alpha = .2) +  # alpha to fade lines
  geom_point() +
  geom_point(aes(y = predicted), shape = 5) +
    # > Color adjustments made here...
  geom_point(aes(color = residuals)) +  # Color mapped here
  scale_color_gradient2(low = "blue", mid = "white", high = "red") +  # Colors to use here
  guides(color = FALSE) +
  # <
  theme_bw()
rp
```

QQPlot Linear Regression inverted and regression

```{r}
gauge <- read_excel("~/Desktop/Quarter12/MATH189/HW4/gauge.xlsx")
model <- lm(gauge$gain ~ gauge$density)
yfit<-model$fitted.values
summary(lm(gauge$gain ~ gauge$density))
plot(gauge$density,gauge$gain,xlab='Density',ylab='Gain',main='Snow Gauge Measurements')
lines(gauge$density, yfit)
```

```{r}
d <- gauge
fit <- lm(gain ~ density, data=d)
head(fit, 20)
summary(fit)
plot(fit)
sprintf("Kurtosis of Residuals: %f", kurtosis(fit))
sprintf("Skewness of Residuals: %f", skewness(fit))
d$predicted <- predict(fit)   # Save the predicted values
d$residuals <- residuals(fit) # Save the residual values

```

```{r}
d %>% select(gain, predicted, residuals) %>% head()
rp <- ggplot(d, aes(x = density, y = gain)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightblue") +  # Plot regression slope
  geom_segment(aes(xend = density, yend = predicted), alpha = .2) +  # alpha to fade lines
  geom_point() +
  geom_point(aes(y = predicted), shape = 5) +
    # > Color adjustments made here...
  geom_point(aes(color = residuals)) +  # Color mapped here
  scale_color_gradient2(low = "blue", mid = "white", high = "red") +  # Colors to use here
  guides(color = FALSE) +
  # <
  theme_bw()
rp
```

Added analysis
```{r}
model <- lm(formula = gauge$gain ~ poly(gauge$density, degree=2, raw=TRUE))
summary(model)
confint(model, level=0.95)
plot(fitted(model),residuals(model),xlab='Fitted',ylab='Residuals',main='Fitted vs Residuals Plot')
plot(gauge$density,gauge$gain,xlab='Density',ylab='Gain',main='Snow Gauge Measurements')

predicted.intervals <- predict(model,data.frame(x=gauge$density),interval='confidence',
                               level=0.95)
lines(gauge$density,predicted.intervals[,1],col='red',lwd=1)
lines(gauge$density,predicted.intervals[,2],col='green',lwd=1)
lines(gauge$density,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
        col=c("red","green","blue"), lwd=2)
```

For means, QQPlot Linear Regression
```{r}
m <- data.frame(newdens,newgain)

fit <- lm(newdens ~ newgain)
head(fit, 20)
summary(fit)
plot(fit)
sprintf("Kurtosis of Residuals: %f", kurtosis(fit))
sprintf("Skewness of Residuals: %f", skewness(fit))
m$predicted <- predict(fit)   # Save the predicted values
m$residuals <- residuals(fit) # Save the residual values

```

```{r}

m %>% select(newdens, predicted, residuals) %>% head()
rp <- ggplot(m, aes(x = newgain, y = newdens)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightblue") +  # Plot regression slope
  geom_segment(aes(xend = newgain, yend = predicted), alpha = .2) +  # alpha to fade lines
  geom_point() +
  geom_point(aes(y = predicted), shape = 5) +
    # > Color adjustments made here...
  geom_point(aes(color = residuals)) +  # Color mapped here
  scale_color_gradient2(low = "blue", mid = "white", high = "red") +  # Colors to use here
  guides(color = FALSE) +
  # <
  theme_bw()
rp
```

Simple linear regression with log transformation of gain
```{r}
model <- lm(gauge$density ~ log(gauge$gain))
yfit<-model$fitted.values
summary(lm(gauge$density ~ log(gauge$gain)))
plot(log(gauge$gain),gauge$density,xlab='Logarithm of Gain',ylab='Density',main='Snow Gauge Measurements')
lines(log(gauge$gain), yfit)
```

FOR WEIGHTED LEAST SQUARE

Trying to fit the model up to a 4 degree polynomial with linear regression
```{r}
w <- 1/store
model1 <- lm(formula = gauge$density ~ poly(gauge$gain, degree=1, raw=TRUE),weights=w)
model2 <- lm(formula = gauge$density ~ poly(gauge$gain, degree=2, raw=TRUE),weights=w)
model3 <- lm(formula = gauge$density ~ poly(gauge$gain, degree=3, raw=TRUE),weights=w)
model4 <- lm(formula = gauge$density ~ poly(gauge$gain, degree=4, raw=TRUE),weights=w)
```

Testing the first degree fit
```{r}
summary(model1)
confint(model1, level=0.95)
plot(fitted(model1),residuals(model1),xlab='Fitted',ylab='Residuals',main='Fitted vs Residuals Plot')
```

Plot first degree fit
```{r}
plot(gauge$gain,gauge$density,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')

predicted.intervals <- predict(model1,data.frame(x=gauge$gain),interval='confidence',
                               level=0.95)
lines(gauge$gain,predicted.intervals[,1],col='red',lwd=1)
lines(gauge$gain,predicted.intervals[,2],col='green',lwd=1)
lines(gauge$gain,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
         col=c("red","green","blue"), lwd=2)

```

Testing the second degree fit
```{r}
summary(model2)
confint(model2, level=0.95)
plot(fitted(model2),residuals(model2),xlab='Fitted',ylab='Residuals',main='Fitted vs Residuals Plot')

```

Plot second degree fit
```{r}
plot(gauge$gain,gauge$density,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')

predicted.intervals <- predict(model2,data.frame(x=gauge$gain),interval='confidence',
                               level=0.95)
lines(gauge$gain,predicted.intervals[,1],col='red',lwd=1)
lines(gauge$gain,predicted.intervals[,2],col='green',lwd=1)
lines(gauge$gain,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
         col=c("red","green","blue"), lwd=2)

```

Testing the third degree fit
```{r}
summary(model3)
confint(model3, level=0.95)
plot(fitted(model3),residuals(model3),xlab='Fitted',ylab='Residuals',main='Fitted vs Residuals Plot')
```

Plot third degree fit
```{r}
plot(gauge$gain,gauge$density,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')
yfit<-model3$fitted.values
lines(gauge$gain, yfit)

```

Testing the fourth degree fit
```{r}
summary(model4)
confint(model4, level=0.95)
plot(fitted(model4),residuals(model4),xlab='Fitted',ylab='Residuals',main='Fitted vs Residuals Plot')
```

Plot fourth degree fit
```{r}
plot(gauge$gain,gauge$density,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')

predicted.intervals <- predict(model4,data.frame(x=gauge$gain),interval='confidence',
                               level=0.95)
lines(gauge$gain,predicted.intervals[,1],col='red',lwd=1)
lines(gauge$gain,predicted.intervals[,2],col='green',lwd=1)
lines(gauge$gain,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
         col=c("red","green","blue"), lwd=2)


```

FOR GROUPED BY MEANS

Trying to fit the model up to a 4 degree polynomial with linear regression
```{r}
model1 <- lm(formula = newdens ~ poly(newgain, degree=1, raw=TRUE))
model2 <- lm(formula = newdens ~ poly(newgain, degree=2, raw=TRUE))
model3 <- lm(formula = newdens ~ poly(newgain, degree=3, raw=TRUE))
model4 <- lm(formula = newdens ~ poly(newgain, degree=4, raw=TRUE))
```

Testing the first degree fit
```{r}
summary(model1)
confint(model1, level=0.95)
plot(fitted(model1),residuals(model1),xlab='Fitted',ylab='Residuals',main='Fitted vs Residuals Plot')
```

Plot first degree fit
```{r}
plot(newgain,newdens,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')

predicted.intervals <- predict(model1,data.frame(x=newgain),interval='confidence',
                               level=0.95)
lines(newgain,predicted.intervals[,1],col='red',lwd=1)
lines(newgain,predicted.intervals[,2],col='green',lwd=1)
lines(newgain,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
         col=c("red","green","blue"), lwd=2)

```

Testing the second degree fit
```{r}
summary(model2)
confint(model2, level=0.95)
plot(fitted(model2),residuals(model2),xlab='Fitted',ylab='Residuals',main='Fitted vs Residuals Plot')
```

Plot second degree fit
```{r}
plot(newgain,newdens,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')
yfit<-model2$fitted.values
lines(newgain, yfit)
```
Testing the third degree fit
```{r}
summary(model3)
confint(model3, level=0.95)
plot(fitted(model3),residuals(model3),xlab='Fitted',ylab='Residuals',main='Fitted vs Residuals Plot')
```

Plot third degree fit
```{r}
plot(newgain,newdens,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')

predicted.intervals <- predict(model3,data.frame(x=newgain),interval='confidence',
                               level=0.95)
lines(newgain,predicted.intervals[,1],col='red',lwd=1)
lines(newgain,predicted.intervals[,2],col='green',lwd=1)
lines(newgain,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
         col=c("red","green","blue"), lwd=2)

```

PART II:

WLS
```{r}
gauge <- read_excel("~/Desktop/Quarter12/MATH189/HW4/gauge.xlsx")
model3 <- lm(formula = gauge$density ~ poly(gauge$gain, degree=3, raw=TRUE),weights=w)

plot(gauge$gain,gauge$density,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')

predicted.intervals <- predict(model3,data.frame(x=gauge$gain),interval='confidence',
                               level=0.95)
lines(gauge$gain,predicted.intervals[,1],col='red',lwd=1)
lines(gauge$gain,predicted.intervals[,2],col='green',lwd=1)
lines(gauge$gain,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
         col=c("red","green","blue"), lwd=2)

```

Means
```{r}
plot(newgain,newdens,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')

model2 <- lm(formula = newdens ~ poly(newgain, degree=2, raw=TRUE))

predicted.intervals <- predict(model2,data.frame(x=newgain),interval='confidence',
                               level=0.95)
lines(newgain,predicted.intervals[,1],col='red',lwd=1)
lines(newgain,predicted.intervals[,2],col='green',lwd=1)
lines(newgain,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
         col=c("red","green","blue"), lwd=2)
```

PART III:

Omitting density 0.508, creating the weights, with variance
```{r}
gauge <- read_excel("~/Desktop/Quarter12/MATH189/HW4/gauge.xlsx")
gauge <- gauge[ ! gauge$density %in% 0.508, ]

d <- gauge$gain
x=0
store=vector()
for (i in 1:8){
  variance=vector()
  for (j in 1:10){
    variance[j] <- d[j+x]
  }
  x=x+10
  store[i] <- var(variance)
}
store <- rep(store,each=10)
```

Omitting density 0.508, creating the means
```{r}
gauge <- read_excel("~/Desktop/Quarter12/MATH189/HW4/gauge.xlsx")
gauge <- gauge[ ! gauge$density %in% 0.508, ]

g <- gauge$gain
x=0
newgain=vector()
for (i in 1:9){
  avg=vector()
  for (j in 1:10){
    avg[j] <- g[j+x]
  }
  x=x+10
  newgain[i] <- mean(avg)
}

d <- gauge$density
y=0
newdens=vector()
for (k in 1:9){
  newdens[k] <- d[k+y]
  y=y+10
}
```

Trying to fit the model up to a 3 degree polynomial with linear regression
```{r}
w <- 1/store
model3 <- lm(formula = gauge$density ~ poly(gauge$gain, degree=3, raw=TRUE),weights=w)
```

Getting the polynomial for the third degree, building the estimate and confidence interval with plot
```{r}
summary(model3)
confint(model3, level=0.95)
plot(gauge$gain,gauge$density,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')

predicted.intervals <- predict(model3,data.frame(x=gauge$gain),interval='confidence',
                               level=0.95)
lines(gauge$gain,predicted.intervals[,1],col='red',lwd=1)
lines(gauge$gain,predicted.intervals[,2],col='green',lwd=1)
lines(gauge$gain,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
         col=c("red","green","blue"), lwd=2)

```

Trying to fit the model up to a 2 degree polynomial with linear regression
```{r}
model2 <- lm(formula = newdens ~ poly(newgain, degree=2, raw=TRUE))
```

Getting the polynomial for the second degree, building the estimate and confidence interval with plot
```{r}
summary(model2)
confint(model2, level=0.95)
plot(newgain,newdens,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')

predicted.intervals <- predict(model2,data.frame(x=newgain),interval='confidence',
                               level=0.95)
lines(newgain,predicted.intervals[,1],col='red',lwd=1)
lines(newgain,predicted.intervals[,2],col='green',lwd=1)
lines(newgain,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
         col=c("red","green","blue"), lwd=2)

```

Omitting density 0.001, creating the weights, with variance
```{r}
gauge <- read_excel("~/Desktop/Quarter12/MATH189/HW4/gauge.xlsx")
gauge <- gauge[ ! gauge$density %in% 0.001, ]

d <- gauge$gain
x=0
store=vector()
for (i in 1:8){
  variance=vector()
  for (j in 1:10){
    variance[j] <- d[j+x]
  }
  x=x+10
  store[i] <- var(variance)
}
store <- rep(store,each=10)
```

Omitting density 0.001, creating the means
```{r}
gauge <- read_excel("~/Desktop/Quarter12/MATH189/HW4/gauge.xlsx")
gauge <- gauge[ ! gauge$density %in% 0.001, ]

g <- gauge$gain
x=0
newgain=vector()
for (i in 1:9){
  avg=vector()
  for (j in 1:10){
    avg[j] <- g[j+x]
  }
  x=x+10
  newgain[i] <- mean(avg)
}

d <- gauge$density
y=0
newdens=vector()
for (k in 1:9){
  newdens[k] <- d[k+y]
  y=y+10
}
```

Trying to fit the model up to a 3 degree polynomial with linear regression
```{r}
w <- 1/store
model3 <- lm(formula = gauge$density ~ poly(gauge$gain, degree=3, raw=TRUE),weights=w)
```

Getting the polynomial for the third degree, building the estimate and confidence interval with plot
```{r}
summary(model3)
confint(model3, level=0.95)
plot(gauge$gain,gauge$density,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')

predicted.intervals <- predict(model3,data.frame(x=gauge$gain),interval='confidence',
                               level=0.95)
lines(gauge$gain,predicted.intervals[,1],col='red',lwd=1)
lines(gauge$gain,predicted.intervals[,2],col='green',lwd=1)
lines(gauge$gain,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
         col=c("red","green","blue"), lwd=2)
```

Trying to fit the model up to a 2 degree polynomial with linear regression
```{r}
model2 <- lm(formula = newdens ~ poly(newgain, degree=2, raw=TRUE))
```

Getting the polynomial for the second degree, building the estimate and confidence interval with plot
```{r}
summary(model2)
confint(model2, level=0.95)
plot(newgain,newdens,xlab='Gain',ylab='Density',main='Snow Gauge Measurements')

predicted.intervals <- predict(model2,data.frame(x=newgain),interval='confidence',
                               level=0.95)
lines(newgain,predicted.intervals[,1],col='red',lwd=1)
lines(newgain,predicted.intervals[,2],col='green',lwd=1)
lines(newgain,predicted.intervals[,3],col='blue',lwd=1)
  legend("topright",c("Fit","Lower","Upper"), 
         col=c("red","green","blue"), lwd=2)

```
